---
title: "Final Report"
author: "Sere Williams", "Heather Deel", "Sunny Liu", "Kuhm Thapa-Magar"
date: "11/27/2019"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research Question

Our research question aims to ask which categories appear most often in the 35 seasons that Jeopardy has run. In this analysis, we will look at the top 5 most repeated categories and themes of the numerous categories that were only used once. To look at category themes, we will use a word search to group category types and isolate commonalities. Lastly, we will use a word search to identify current events and link the event to the season inwhich that category was used. 

## Loading libraries

```{r}
# Khum did..
# loading libraries for whole analysis
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(readr)
```



## Read in the data
```{r}
seasons_all <- read_tsv("master_season1-35.tsv")
```

## Create new dataframe with all seasons
```{r}
seasons_all_cat <- seasons_all %>%
  group_by(category) %>% 
  count() %>% 
  arrange(desc(n))
```




