---
title: "Flexdashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
---

Row 
-------------------------------------

```{r setup, include=FALSE}
library(flexdashboard)
```

## Research questions:
Question 1: What are the top 5 categories in all seasons, and how often do these occur per year?
Question 2: What are the top 5 categories that different kinds of gods (Greek, Norse, Hindu) are mentioned either in the question or answer?
How often to the two popular Greek gods, Zeus and Athena, appear over time and in the top 5 categories discovered in question 2?

# Loading libraries

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(readr)
library(lubridate)
```

# Read in and clean the data


```{r}
seasons_all <- read_tsv("master_season1-35.tsv")
seasons_all_clean <- seasons_all %>% 
  separate(air_date, sep="-", into = c("year", "month", "day")) %>% 
  select(-comments, -notes) %>% 
  mutate(category = str_to_lower(string = category)) %>%
  mutate(answer = str_to_lower(string = answer)) %>% 
  mutate(question = str_to_lower(string = question)) %>% 
  filter(year != "2019")
```

## Question 1: What are the top 5 categories in all seasons, and how often do these occur per year?
######################################
######### SERE PUT CODE HERE
######################################

## Question 2: What are the top 5 categories that different kinds of gods (Greek, Norse, Hindu) are mentioned either in the question or answer?

# Create dataframe filtered by different kinds of gods, all merged into one dataframe
```{r}
# Create lists of gods
greek_gods <- c("athena", "zeus", "poseidon", "hades", "hera", "apollo", "demeter", "artemis", "aphrodite", "ares")
norse_gods <- c("odin", "loki", "thor", "njord", "heimdall", "frigg", "tyr", "baldur", "idun", "bragi")
hindu_gods <- c("vishnu", "brahma", "shiva", "lakshmi", "hanuman", "ganesha", "krishna", "kali", "rama", "saraswati")

# Create filtered datasets for each type of god
seasons_all_greek <- seasons_all_clean %>% 
  filter(str_detect(question, greek_gods) | str_detect(answer, greek_gods))

seasons_all_norse <- seasons_all_clean %>% 
  filter(str_detect(question, norse_gods) | str_detect(answer, norse_gods))

seasons_all_hindu <- seasons_all_clean %>% 
  filter(str_detect(question, hindu_gods) | str_detect(answer, hindu_gods))

# Merge all 3 dataframes
seasons_all_gods <- merge(merge(seasons_all_greek, seasons_all_norse, all = TRUE), seasons_all_hindu, all = TRUE)
```
# Group by category, filter by top 5 categories, and graph

```{r}
seasons_all_gods_cat <- seasons_all_gods %>% 
  group_by(category) %>% 
  count()

seasons_all_gods_top5 <- seasons_all_gods_cat %>%
    filter(category == "authors" | 
         category == "books & authors" |
         category == "literature" | 
         category == "mythology" | 
         category == "english literature")

seasons_all_gods_top5 %>%
  ggplot(aes(x = reorder(category, -n), y = n)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme_classic() +
  ggtitle("Number of times Greek, Norse, and Hindu Gods are mentioned") +
  labs(x = "Category", y = "Number of mentions")
```

## Question 3: How often to the two popular Greek gods, Zeus and Athena, appear over time and in the top 5 categories discovered in question 2?

# Create a god_name column that specifies whether Zeus or Athena was mentioned in the question or answer
```{r}
seasons_all_gods_athenaQ <- seasons_all_gods %>% 
  mutate(god_name = stringr::str_extract(question, "athena"))

seasons_all_gods_athenaA <- seasons_all_gods %>% 
  mutate(god_name = stringr::str_extract(answer, "athena"))

seasons_all_gods_zeusQ <- seasons_all_gods %>% 
  mutate(god_name = stringr::str_extract(question, "zeus"))

seasons_all_gods_zeusA <- seasons_all_gods %>% 
  mutate(god_name = stringr::str_extract(answer, "zeus"))

seasons_all_gods_ZA <- merge(merge(merge(seasons_all_gods_athenaQ, seasons_all_gods_athenaA, all = TRUE), seasons_all_gods_zeusQ, all = TRUE), seasons_all_gods_zeusA, all = TRUE)
```

# Fill by Zeus or Athena to look at their prevalences over time
```{r}
seasons_all_gods_ZA_group <- seasons_all_gods_ZA %>% 
  group_by(category, year, god_name) %>% 
  count()

seasons_all_gods_ZA_group %>% 
  ggplot(aes(x = year, y = n, fill = god_name)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_classic() +
  ggtitle("The prevalence of Zeus and Athena over time") +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(x = "Year", y = "Number of mentions", fill = "God Name")

### note that this is testing the prevalence of Zeus and Athena out of all 15 tested gods from all categories
```

# Filter the ZA dataframe to top 5 categories and fill by Zeus or Athena
```{r}
seasons_all_gods_ZA_top5 <- seasons_all_gods_ZA %>%
      filter(category == "authors" |
         category == "books & authors" |
         category == "literature" | 
         category == "mythology" | 
         category == "english literature")

seasons_all_gods_ZA_top5_group <- seasons_all_gods_ZA_top5 %>% 
  group_by(category, year, god_name) %>% 
  count()

seasons_all_gods_ZA_top5_group %>%
  ggplot(aes(x = reorder(category, -n), y = n, fill = god_name)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  ggtitle("Number of times Zeus and Athena are mentioned in the most popular categories") +
  labs(x = "Category", y = "Number of mentions", fill = "God Name")
```