---
title: "Jeopardy Project All Seasons"
author: "Heather Deel"
date: "11/18/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(forcats)
library(RColorBrewer)
```

## Read in the data
```{r}
seasons_all <- read_tsv("master_season1-35.tsv")
```

## Create new dataframe with all seasons
```{r}
seasons_all_cat <- seasons_all %>%
  group_by(category) %>% 
  count() %>% 
  arrange(desc(n))
```

## Clean up seasons data
```{r}
seasons_all_clean <- seasons_all %>% 
  separate(air_date, sep="-", into = c("year", "month", "day")) %>% 
  select(-comments, -notes) %>% 
  mutate(category = str_to_lower(string = category)) %>%
  mutate(answer = str_to_lower(string = answer)) %>% 
  mutate(question = str_to_lower(string = question)) %>% 
  filter(year != "2019")
```

Question 2: How often does the mythology category appear over each year? Is there a trend in how often it occurs over time?

## Filter by mythology category and group by year
```{r}
seasons_all_myth <- seasons_all_clean %>% 
  filter(category == "mythology") %>% 
  group_by(year) %>% 
  count()
```

## Plot mythology category by year
```{r}
seasons_all_myth %>% 
  ggplot(aes(x = year, y = n)) +
  geom_bar(stat = "identity", fill = "pink") +
  theme_classic() +
  ggtitle("Number of questions asked in the Mythology category by season") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Year", y = "Number of questions")
  
```

Question 3: How often does “Athena”, the greek goddess of war, appear over time? Is there a trend?

## Filter to "Athena" in question and answer columns and group by year
```{r}

# note that with this code, it includes "Athena" not just in the mythology category - increases occurrences from 21 to 103

seasons_all_athena <- seasons_all_clean %>% 
  filter(str_detect(question, "athena") | str_detect(answer, "athena")) %>% 
  group_by(year) %>% 
  count()

# create dataframe without grouping to see what other categories are there
seasons_all_athena_cat <- seasons_all_clean %>% 
  filter(str_detect(question, "athena") | str_detect(answer, "athena")) %>% 
  group_by(category) %>% 
  count()
```

## Plot "Athena" by season
```{r}
seasons_all_athena %>% 
  ggplot(aes(x = year, y = n)) +
  geom_bar(stat = "identity", fill = "purple") +
  theme_classic() +
  ggtitle("Number of times 'Athena' is mentioned in the question or answer") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Year", y = "Number of mentions")
  
```

## Try plotting but by category
```{r}
seasons_all_athena_cat %>% 
  ggplot(aes(x = category, y = n)) +
  geom_bar(stat = "identity", fill = "purple") +
  theme_classic() +
  ggtitle("Number of times 'Athena' is mentioned in the question or answer") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Category", y = "Number of mentions")
```


### Do it with multiple greek gods

## Filter to multiple greek gods in question and answer columns and group by year
```{r}

greek_gods <- c("athena", "zeus", "poseidon", "hades", "hera", "apollo", "demeter", "artemis", "aphrodite", "ares")

seasons_all_greek <- seasons_all_clean %>% 
  filter(str_detect(question, greek_gods) | str_detect(answer, greek_gods))

```

## Norse
```{r}
norse_gods <- c("odin", "loki", "thor", "njord", "heimdall", "frigg", "tyr", "baldur", "idun", "bragi")

seasons_all_norse <- seasons_all_clean %>% 
  filter(str_detect(question, norse_gods) | str_detect(answer, norse_gods))
```

## Graph norse gods by category
```{r}
seasons_all_norse_top5 <- seasons_all_norse %>% 
  filter(category == "authors" |
         category == "books & authors" |
         category == "literature" | 
         category == "english literature" | 
         category == "fictional characters") %>% 
  group_by(category) %>% 
  count()
  

seasons_all_norse_top5 %>%
  ggplot(aes(x = reorder(category, -n), y = n)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme_classic() +
  ggtitle("Number of times Norse Gods are mentioned in the question or answer") +
  labs(x = "Category", y = "Number of mentions")
```


## Hindu
```{r}
hindu_gods <- c("vishnu", "brahma", "shiva", "lakshmi", "hanuman", "ganesha", "krishna", "kali", "rama", "saraswati")

seasons_all_hindu <- seasons_all_clean %>% 
  filter(str_detect(question, hindu_gods) | str_detect(answer, hindu_gods))
```

### Try adding column for god name
```{r}
seasons_all_hindu_name <- seasons_all_clean %>% 
  filter(str_detect(question, hindu_gods) | str_detect(answer, hindu_gods)) %>% 
  mutate(god_name = stringr::str_extract(question, "vishnu"))
  
  
```


## Merga all 3 datasets
```{r}
seasons_all_gods3 <- merge(merge(seasons_all_greek, seasons_all_norse, all = TRUE), seasons_all_hindu, all = TRUE)
```

## Group by category
```{r}
seasons_all_gods_cat <- seasons_all_gods3 %>% 
  group_by(category) %>% 
  count()

seasons_all_gods_top5 <- seasons_all_gods_cat %>%
    filter(category == "authors" |
         category == "books & authors" |
         category == "literature" | 
         category == "mythology" | 
         category == "english literature")
  
```

## Now graph
```{r}
seasons_all_gods_top5 %>%
  ggplot(aes(x = reorder(category, -n), y = n)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme_classic() +
  ggtitle("Number of times all Gods are mentioned in the question or answer") +
  labs(x = "Category", y = "Number of mentions")
```

### will maybe try to facet by god or something for another interesting figure

### In depth analysis of gods and their prevalence in certain categories

### create columns for Zeus and Athena

```{r}
seasons_all_gods3_athenaQ <- seasons_all_gods3 %>% 
  mutate(god_name = stringr::str_extract(question, "athena"))

seasons_all_gods3_athenaA <- seasons_all_gods3 %>% 
  mutate(god_name = stringr::str_extract(answer, "athena"))

seasons_all_gods3_zeusQ <- seasons_all_gods3 %>% 
  mutate(god_name = stringr::str_extract(question, "zeus"))

seasons_all_gods3_zeusA <- seasons_all_gods3 %>% 
  mutate(god_name = stringr::str_extract(answer, "zeus"))

seasons_all_gods_ZAcat <- merge(merge(merge(seasons_all_gods3_athenaQ, seasons_all_gods3_athenaA, all = TRUE), seasons_all_gods3_zeusQ, all = TRUE), seasons_all_gods3_zeusA, all = TRUE)

```


### Fill by god (Zeus or Athena) to look at their prevalences over time
```{r}
seasons_all_gods_ZAcat_group <- seasons_all_gods_ZAcat %>%
  group_by(category, year, god_name) %>% 
  count()
```

## Releveling
```{r}
# make it so it says 'other'
seasons_all_gods_ZAcat$god_name <- as.character(seasons_all_gods_ZAcat$god_name)
seasons_all_gods_ZAcat$god_name[is.na(seasons_all_gods_ZAcat$god_name)] <- "other"

seasons_all_gods_ZAcat$god_name <- as.factor(seasons_all_gods_ZAcat$god_name)

seasons_all_gods_ZAcat$god_name <- relevel(seasons_all_gods_ZAcat$god_name, "other")
```


```{r}

color_pal <- c("grey75", "tomato1", "royalblue1")

seasons_all_gods_ZAcat_group %>%
  ggplot(aes(x = year, y = n, fill = god_name)) +
  scale_fill_manual(values = color_pal) +
  geom_bar(position = "stack", stat = "identity") +
  theme_classic() +
  ggtitle("The prevalence of Zeus and Athena in questions and answers over time") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

### note that this is testing the prevalence of Zeus and Athena out of all 15 tested gods from all categories

```


### Then filter to the top 5 categories (authors, books & authors, literature, mythology, english literature, and fill by god)

```{r}
seasons_all_gods_ZAcat_top5 <- seasons_all_gods_ZAcat %>%
      filter(category == "authors" |
         category == "books & authors" |
         category == "literature" | 
         category == "mythology" | 
         category == "english literature")

seasons_all_gods_ZAcat_top5_group <- seasons_all_gods_ZAcat_top5 %>% 
  group_by(category, year, god_name) %>% 
  count()
  
```


```{r}
seasons_all_gods_ZAcat_top5_group %>%
  ggplot(aes(x = reorder(category, -n), y = n, fill = god_name)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  ggtitle("Number of times all Gods are mentioned in the question or answer") +
  labs(x = "Category", y = "Number of mentions")


### use forcats to change 'NA' to 'other and reorder so that Zeus and Athena are on the bottom of the chart

```




