---
title: "SAW_script_all_seasons"
author: "Sere Williams"
date: "11/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research Question

Our research question aims to ask which categories appear most often in the 35 seasons that Jeopardy has run. In this analysis, we will look at the top 5 most repeated categories and themes of the numerous categories that were only used once. To look at category themes, we will use a word search to group category types and isolate commonalities. Lastly, we will use a word search to identify current events and link the event to the season inwhich that category was used. 

## Libraries
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(ggplot2)
```


## Read in the data
```{r}
seasons_all <- read_tsv("master_season1-35.tsv")
```


## Create new dataframe with all seasons
```{r}
seasons_all_cat <- seasons_all %>%
  group_by(category) %>% 
  count() %>% 
  arrange(desc(n))
```


## All seasons cleaned data
```{r}
seasons_all_clean <- seasons_all %>% 
  separate(air_date, sep="-", into = c("year", "month", "day")) %>%
  mutate(year = as.numeric(year)) %>% 
  select(-comments, -notes) %>% 
  mutate(category = str_to_lower(string = category),
         answer = str_to_lower(string = answer),
         question = str_to_lower(string = question)) %>% 
  filter(year != "2019")

```


## Number of questions asked per season
```{r}
seasons_all_count <- seasons_all_clean %>% 
  group_by(year) %>% 
  count()
```



## Visualize top five categories over all seasons
We want to look at the distribution of repeated categories by season.
```{r}
seasons_all_cat_top5 <- seasons_all_clean %>% 
  filter(category == "science" |
         category == "literature" |
         category == "american history" |
         category == "history" |
         category == "sports")

top5_year_total <- seasons_all_cat_top5 %>% 
  group_by(year) %>% 
  count()
```



## Question counts by category by year
```{r}
top5_year_count <- seasons_all_cat_top5 %>% 
  select(-answer, -question, -month, -day) %>% 
  group_by(category, year) %>% 
  tally() %>% 
  arrange(year)
  
```


## Plot the question counts by category by year
Source for stacked bar charts: https://www.r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html
```{r fig.width = 8, fig.height = 4}
top5_year_count_plot <- top5_year_count %>% 
  ggplot(aes(x = year, y = n, fill = category)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_bw() +
  ggtitle("Counts of questions asked in top 5 categories by season") +
  theme(axis.text.x = element_text(angle = 45))

top5_year_count_plot

```


## Counts of questions fasceted by category
```{r}
top5_year_count_plot <- top5_year_count %>% 
  ggplot(aes(x = year, y = n)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ category) +
  theme_bw() +
  ggtitle("Counts of questions asked in top 5 categories by season") +
  theme(axis.text.x = element_text(angle = 45))

top5_year_count_plot
```



## Proportion of top5 categories from all questions by year
```{r}

prop_top5 <- top5_year_count %>% 
  full_join(seasons_all_count, top5_year_count, by = "year") %>% 
  rename("count" = "n.x") %>% 
  rename("total" = "n.y") %>% 
  mutate(proportion = (count/total)*1000)

prop_top5_plot <- prop_top5 %>%
  ggplot(aes(x = year, y = proportion, fill = category)) +
  geom_bar(position = "stack", stat = "identity") +
  theme_bw() +
  ggtitle("Proportion of questions asked from top 5 categories (x1000) by season") +
  theme(axis.text.x = element_text(angle = 45))

prop_top5_plot




```



## What were the top 5 most commonly asked categories from 1998 on?
```{r}
seasons_1998on_count <- seasons_all_clean %>% 
  filter(year >= 1998) %>% 
  group_by(category) %>% 
  count() %>% 
  arrange(desc(n))

seasons_1998on_top5 <- seasons_all_clean %>% 
  filter(year >= 1998) %>% 
  filter(category == "before & after" |
         category == "word origins" |
         category == "american history" |
         category == "potpourri" |
         category == "literature")

tally_1998on <- seasons_1998on_top5 %>% 
  select(-answer, -question, -month, -day) %>% 
  group_by(category, year) %>% 
  tally() %>% 
  arrange(year)

tally_1998on_plot <- tally_1998on %>% 
  ggplot(aes(x = year, y = n)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ category) +
  theme_bw() +
  ggtitle("Counts of questions asked in top 5 categories by season from 1998 on") +
  theme(axis.text.x = element_text(angle = 45))

tally_1998on_plot
```










